---
title: "PBMC Flex"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(Seurat)
library(BPCells) #need to install via github
library(tidyverse)
library(AzimuthAPI) # need to install via github with devtools or remotes
# needs to be set for large dataset analysis
options(future.globals.maxSize = 5e9)

# get a list of all the .h5 files exported from CellRanger
files <- list.files("data/input", recursive = TRUE, 
  full.names = TRUE) %>% 
  # ONLY include the "sample_filtered_feature_bc_matrix.h5" files, 
  # ignore all others
  .[str_detect(., "sample_filtered_feature_bc_matrix.h5")]

# convert file names vector to a dataframe containing the full file names
# as well as the corresponding GEM and Sample name
files_df <- tibble(Files = files) %>% 
  mutate(GEM = str_extract(Files, "GEM.")) %>% 
  mutate(Sample = str_extract(Files, "MP-\\d\\d*"))
```

```{r functions}
# THIS CHUNK INCLUDES FUNCTIONS TO READ IN THE RAW FILES

# read in one h5 file, create a Seurat object, add metadata
# the function takes a single row of a data frame
#     the data frame should include the file name, the GEM, and the Sample
# add GEM and Sample to the meta.data of the resulting Seurat object
read_one_file <- function(files_df_row) {
  one_file <- Read10X_h5(files_df_row$Files) %>% CreateSeuratObject()
  # add metadata
  one_file$GEM <- files_df_row$GEM
  one_file$Sample <- files_df_row$Sample
  
  one_file
}

# use the BPCells package to put a Seurat object on disk to save memory
put_on_disk <- function(obj) {
  sample_id <- obj@meta.data$Sample[1]
  
  write_matrix_dir(mat = obj[["RNA"]]$counts, dir = paste0('data/temp/', sample_id))
  counts.mat <- open_matrix_dir(dir = paste0('data/temp/', sample_id))
  obj[["RNA"]]$counts <- counts.mat
  
  obj  
}

# read and combine all four GEM files for 1 sample
# Each sample is split across the 4 GEMs
# Given a sample name (e.g. "MP-1"), this function reads the 
# 4 h5 files for the sample (from each GEM) and combines them
# Then it puts the object on disk to save memory
read_one_sample <- function(sample) {
  # filter for the files just for one sample
  one_sample <- files_df %>% 
    filter(Sample == sample)
  
  # read in all 4 GEMs for one sample
  one_sample_list <- list(
    read_one_file(one_sample[1,]),
    read_one_file(one_sample[2,]),
    read_one_file(one_sample[3,]),
    read_one_file(one_sample[4,])
  )
  
  # merge together the four samples
  one_sample_merged <- merge(
                        # select first list element
                        one_sample_list[[1]],
                        # merge with every element except the first
                        one_sample_list[-1], 
                        add.cell.ids = c("GEM1", "GEM2", "GEM3", "GEM4") 
                       )

  
  # JoinLayers converts from one "counts layer" per GEM
  # to one per sample
  one_sample_merged %>% 
    JoinLayers() %>% 
    put_on_disk()
}

# test code to make sure each function works
# sample_1 <- read_one_sample("MP-1")
# sample_2 <- read_one_sample("MP-2")
# test <- merge(sample_1, sample_2, add.cell.ids = c("MP-1", "MP-2"))
```

```{r}
# THIS CHUNK READS IN THE RAW FILES AND MERGES THEM INTO A SEURAT OBJECT

# Read in every sample
# This results in a list of seurat objects (one per sample)
all_samples_list <- lapply(unique(files_df$Sample), read_one_sample)

# save list of seurat objects
write_rds(all_samples_list, "data/rds_files/01-list_of_samples.rds")

#read in the saved list of files, not normed yet
# all_samples_list<-read_rds("data/rds_files/01-list_of_samples.rds")

```

```{r}
# THIS CHUNK MERGES THE LIST OF SEURAT OBJECTS INTO A SINGLE OBJECT

# Combine list of seurat objects into a single Seurat object
all_samples_merged <- merge(
                    # select first list element
                    all_samples_list[[1]],
                    # merge with every element except the first
                    all_samples_list[-1], 
  add.cell.ids =  unique(files_df$Sample)
                        )
```

```{r}
# THIS CHUNK ADDS METADATA (PTID AND PHASE) TO THE SAMPLES

# To add metadata with AddMetaData, you need a dataframe with rownames that match 
# the rownames from the SeuratObject 
# Steps:
# 1. Load metadata from file
metadata <- readxl::read_excel("data/PBMC-multilab plan for Innovation Lab_9feb26.xlsx", sheet = "Prlic") %>% 
    mutate(Phase = str_sub(`Sample ID`, 8, 8),
        Ptid = str_sub(`Sample ID`, 1, 7))

# 2. Select columns you want to add 
to_add <- metadata %>% 
    select(Sample = `Tube sample`, Ptid, Phase) %>% 
    as.data.frame()

# 3. Get the metadata out of the Seurat object because it has the rownames
old_meta_data <- all_samples_merged@meta.data %>% 
    mutate(Rownames = rownames(.))

# 4. Merge the two metadata objects by sample. This associates all your new metadata
# with the appropriate cell identifiers in the rownames. Make sure to keep the rownames
# which tidyverse removes
new_meta_data <- merge(old_meta_data, to_add, by = "Sample")
rownames(new_meta_data) <- new_meta_data$Rownames
new_meta_data <- select(new_meta_data, -Rownames)

# 5. Add the new metadata to the Seurat Object
all_samples_merged <- AddMetaData(all_samples_merged, new_meta_data)
```

```{r}
# THIS CHUNK CHECKS THE QC METRICS AND FILTERS OUT LOW QUALITY CELLS

# check QC metrics

# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
all_samples_merged[["percent.mt"]] <- PercentageFeatureSet(all_samples_merged, pattern = "^MT-")

# Visualize QC metrics as a violin plot
# have some cells with  really high transcript counts, most cells have info about ~2k genes, a few cells have a higher percentage of mt transcripts

VlnPlot(all_samples_merged, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)

VlnPlot(all_samples_merged, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, group.by = "GEM")
ggsave("images/violins_by_gem.png", width = 6, height = 4, dpi = 600)

VlnPlot(all_samples_merged, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, group.by = "Phase")
ggsave("images/violins_by_phase.png", width = 6, height = 4, dpi = 600)

VlnPlot(all_samples_merged, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, group.by = "Ptid")
ggsave("images/violins_by_ptid.png", width = 8, height = 4, dpi = 600)

#dropoff at about 20k
all_samples_merged$nCount_RNA  %>%  
    hist(breaks= 100)

#lowest counts are around 450
sort(all_samples_merged$nCount_RNA) %>% tail(n = 20)
#highest are around 140k
sort(all_samples_merged$nCount_RNA) %>% head(n = 20)

all_samples_merged$nFeature_RNA  %>%  
    hist(breaks= 100)

#highest is around 8500
sort(all_samples_merged$nFeature_RNA) %>% tail(n = 20)
#lowest are around 40
sort(all_samples_merged$nFeature_RNA) %>% head(n = 20)


# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
plot1 <- FeatureScatter(all_samples_merged, feature1 = "nCount_RNA", feature2 = "percent.mt")


plot2 <- FeatureScatter(all_samples_merged, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2



#subset for nCount, nFeature and percent.mt. Tutorial does not have an nCount subset but we are adding one because we don't want info from the cells that only have a few transcripts read
#our cutoffs are based on the histograms made above

all_samples_merged_subset <- subset(all_samples_merged, 
    subset = nFeature_RNA > 200 & nFeature_RNA < 6000 & percent.mt < 10 & nCount_RNA < 2E4)

# save merged seurat object
write_rds(all_samples_merged_subset, "data/rds_files/02-merged-subset.rds")

#check violins again after subsetting
VlnPlot(all_samples_merged_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)

VlnPlot(all_samples_merged_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, group.by = "GEM")
ggsave("images/violins_by_gem_subset.png", width = 6, height = 4, dpi = 600)

VlnPlot(all_samples_merged_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, group.by = "Phase")
ggsave("images/violins_by_phase_subset.png", width = 6, height = 4, dpi = 600)

VlnPlot(all_samples_merged_subset, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, group.by = "Ptid")
ggsave("images/violins_by_ptid_subset.png", width = 8, height = 4, dpi = 600)


#comparing our original and subset objects, we've removed about 5k cells, not much!! Lol
all_samples_merged_subset
all_samples_merged
```



```{r}
# THIS CHUNK RUNS SCTRANSFORM

#instead of normalizeData we're trying the cooler sctransform which is better because...it accounts for sequencing depth?
# we are not regressing out any of our variables, we aren't sure if they are important. We ran out of memory on Sean's 16gb machine doing this.
#Running future::plan("sequential" didn't help)

# needs to be set for large dataset analysis, SCTransform did not work with this set to 1E9
options(future.globals.maxSize = 5e9)

#all_samples_merged_subset_norm<-SCTransform(all_samples_merged_subset,
#    vars.to.regress = NULL, verbose = TRUE, conserve.memory = TRUE)

# ended up failing after 5 samples (I think) with this error: 
# Error: cannot allocate vector of size 72.1 Mb
# I think that means it ran out of RAM
# I'll try again tonight after removing the other objects from memory and closing all other programs


#This is what we did the first time for normalization
all_samples_merged_subset <- NormalizeData(all_samples_merged_subset, normalization.method = "LogNormalize", scale.factor = 10000)

# save merged seurat object
write_rds(all_samples_merged_subset, "data/rds_files/02-merged.rds")
```

```{r}
# Read in merged seurat object

all_samples_merged_subset <- read_rds("data/rds_files/02-merged.rds")
```

```{r}
# THIS CHUNK CREATES A SKETCH AND DOES SOME PROCESSING

# Need to find variable features before creating sketch
all_samples_merged_subset <- FindVariableFeatures(all_samples_merged_subset, layer = "data")

# create a "sketch" containing `ncells` cells per sample
# this takes a long time
sketch <- SketchData(
  object = all_samples_merged_subset,
  ncells = 1000,
  method = "LeverageScore",
  sketched.assay = "sketch",
  # see comment below explaining this next line 
  features = VariableFeatures(all_samples_merged_subset)
)

# we were getting this warning/error: "No variable features were found in data.1. Falling back to variable features from counts.1"
# In this link, it says if you have more than 5000 features total, it triggers that warning
# https://github.com/satijalab/seurat/issues/7171
# To remove the warning, add the line specifying which features are variable
# but then you get a new warning: LeverageScore failed on full matrix, retrying with same data (no zero-variance features found).

# save sketch
write_rds(sketch, "data/rds_files/03-sketch-raw.rds")

# switch to analyzing the full dataset (on-disk)
DefaultAssay(sketch) <- "RNA"
# switch to analyzing the sketched dataset (in-memory)
DefaultAssay(sketch) <- "sketch"

# do a bunch of processing on the sketch
sketch <- FindVariableFeatures(sketch)
sketch <- ScaleData(sketch)
sketch <- RunPCA(sketch)
sketch <- FindNeighbors(sketch, dims = 1:30)
sketch <- FindClusters(sketch, resolution = 2)
sketch <- RunUMAP(sketch, reduction = "pca", dims = 1:30, return.model = T, reduction.name = "umap")

# save processed sketch
write_rds(sketch, "data/rds_files/04-sketch-final.rds")
```

```{r read in processed sketch}
sketch<-read_rds("data/rds_files/04-sketch-final.rds")
```

```{r}
# THIS CHUNK PLOTS UMAPS OF THE SKETCH COLORED BY VARIOUS CO-VARIATES

DimPlot(sketch, label = T, label.size = 3, reduction = "umap", group.by = "Sample") + NoLegend()
ggsave("images/umap_by_sample.png", width = 6, height = 4, dpi = 600)

DimPlot(sketch, label = T, label.size = 3, reduction = "umap", group.by = "Phase") + NoLegend()
ggsave("images/umap_by_phase.png", width = 6, height = 4, dpi = 600)

DimPlot(sketch, label = T, label.size = 3, reduction = "umap", group.by = "Ptid") + NoLegend()
ggsave("images/umap_by_ptid.png", width = 6, height = 4, dpi = 600)

DimPlot(sketch, label = T, label.size = 3, reduction = "umap", group.by = "GEM") + NoLegend()
ggsave("images/umap_by_gem.png", width = 6, height = 4, dpi = 600)

# Tried to facet by Ptid with "split.by" argument of DimPlot, but it was really slow
# instead here extract the UMAP from the Seurat object and the metadata and plot separately
umap <- as.data.frame(Embeddings(sketch, reduction = "umap")) %>%
    merge(sketch@meta.data, by = "row.names")

ggplot(umap, aes(umap_1, umap_2, color = seurat_clusters)) + facet_wrap(~Ptid) + geom_point() + ylim(0, 2.5) + xlim(-2.5, 0)
```

## Cluster annotation

```{r}
# THIS CHUNK DOES AZIMUTH CELL ANNOTATION

# This code works starting from 04-sketch-final (the object before running ProjectData)
# but fails after running ProjectData
# See: https://github.com/satijalab/AzimuthAPI/issues/10 for possible solution

# For demonstration, run on a subset of the cells
# Select only those cells that have a seurat cluster (the "sketch" cells)
for_azimuth <- subset(sketch, !is.na(seurat_clusters))


# Have to run JoinLayers before Azimuth or you get an error 
for_azimuth <- JoinLayers(for_azimuth)

# I don't know why but assay has to be "sketch" otherwise it breaks with 
# a very obscure error message
for_azimuth <- AzimuthAPI::CloudAzimuth(for_azimuth, assay = "sketch")

head(for_azimuth@meta.data)

#how many occurrances of each label are there
for_azimuth@meta.data %>% count(azimuth_label) %>% arrange(desc(n))


DimPlot(for_azimuth, group.by = "azimuth_label", label.size = 4, label = T, reduction = "umap",
    repel = TRUE) + NoLegend() + 
    ggtitle("Azimuth labels on un-integrated UMAP")
ggsave("images/umap_by_azimuth_no_integration.png", width = 10, height = 6, dpi = 600)

for_azimuth@meta.data %>% ggplot(aes(x = azimuth_label, fill = factor(seurat_clusters))) + geom_bar() + coord_flip()

for_azimuth@meta.data %>% ggplot(aes(x = seurat_clusters, fill = factor(azimuth_label))) + geom_bar() + coord_flip()

for_azimuth@meta.data %>% filter(str_detect(azimuth_label, "T cell")) %>%  ggplot(aes(x = seurat_clusters, fill = factor(azimuth_label))) + geom_bar() + coord_flip()

DimPlot(for_azimuth, label = T, label.size = 3, reduction = 'umap', group.by = 'azimuth_medium') + NoLegend()



DimPlot(for_azimuth, label = T, label.size = 3, reduction = 'umap', group.by = 'azimuth_medium') + NoLegend()


for_azimuth@meta.data$azimuth_medium%>% unique()

write_rds(for_azimuth, "data/rds_files/for_azimuth.rds")


```

Groups
* NK: 4, 14
* T cell CD4: 0, 1, 3, 9, 10, 13, 18, 19, 20, 23, 24, 25, 26, 30, 31, 34
* T cell CD8: 2, 8, 11, 14, 16, 19, 24, 31 
* B cells: 5, 17, 19, 21, 35
* Monocytes:  6, 15, 20, 23, 25, 28
* 12, 26: platelets
* Multiplets? CD3, CD14, NKG7: 7, 13
* 29: basophils/mast cells
* 30: pDC
* 32: CD34
* 33

```{r}
# THIS CHUNK IDENTIFIES THE MARKERS ASSOCIATED WITH EACH CLUSTER

# using the sketch instead of the projected to make it go faster

sketch <- JoinLayers(sketch)
all_markers <- FindAllMarkers(sketch, only.pos = TRUE)

DotPlot(sketch, features = c("CD3D", "CD4", "CD8A", "NKG7", "CD14", "CD19"))

# are any clusters obviously junk? 
FeaturePlot(sketch, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"))

```

## Project calculations back to full data set

```{r}
# THIS CHUNK PROJECTS THE CALCULATIONS FROM THE SKETCH

# TODO
# Figure out how to project azimuth stuff
# test_project <- ProjectData(
#   object = x,
#   assay = 'RNA',
#   full.reduction = 'pca.full',
#   sketched.assay = 'sketch',
#   sketched.reduction = 'pca',
#   umap.model = 'umap',
#   dims = 1:30,
#   refdata = list(cluster_full = 'seurat_clusters', azimuth_medium_full = "azimuth_medium"))
# pca.full is not in the object. Data from all cells will be projected to pca
# Projecting cell embeddings
# Finding sketch neighbors
# Finding sketch weight matrix
# 0%   10   20   30   40   50   60   70   80   90   100%
# [----|----|----|----|----|----|----|----|----|----|
# **************************************************|
# Transfering refdata from sketch
# Error in TransferLablesNN(reference.labels = reference.labels, weight.matrix = full_sketch.weight) : 
#   wrong weights matrix input
# In addition: Warning message:
# In grid.Call.graphics(C_upviewport, as.integer(n)) :
#   cannot pop the top-level viewport ('grid' and 'graphics' output mixed?)
# x <- TransferSketchLabels(
#   sketch_with_azimuth,
#   sketched.assay = "sketch",
#   reduction = "pca",
#   dims = 1:30,
#   refdata = list(azimuth_medium_full = "azimuth_medium"),
#   k = 50,
#   reduction.model = NULL,
#   neighbors = NULL,
#   recompute.neighbors = FALSE,
#   recompute.weights = FALSE,
#   verbose = TRUE
# )
# same error

projected <- ProjectData(
  object = sketch,
  assay = 'RNA',
  full.reduction = 'pca.full',
  sketched.assay = 'sketch',
  sketched.reduction = 'pca',
  umap.model = 'umap',
  dims = 1:30,
  refdata = list(cluster_full = 'seurat_clusters'))

# now that we have projected the full dataset, switch back to analyzing all cells
DefaultAssay(projected) <- 'RNA'
```

## Add our own annotations based on azimuth and our brains

```{r}
# THIS CHUNK MAPS EACH CELL CLUSTER TO AN ANNOTATION BASED ON AZIMUTH AND OUR
# OBSERVATIONS

# describe each cluster with the most likely azimuth name
cell_annotations <- for_azimuth@meta.data |> 
    # simplify T cell labels to CD4 or CD8 (dropping memory labels)
    mutate(final_level_labels = case_when(
        str_detect(final_level_labels, "CD4") ~ "CD4 T cell", 
        str_detect(final_level_labels, "CD8") ~ "CD8 T cell",
        TRUE ~ final_level_labels
    )) |> 
    # count number of cells with each label in each cluster
    group_by(seurat_clusters, final_level_labels) |> 
    count() |> 
    # for each cluster, use the most common label and count 
    # the percent of cells with that label
    group_by(seurat_clusters) |> 
    arrange(desc(n)) |>
    mutate(percent = n/sum(n)*100) |> 
    slice(1) |> 
    arrange(percent) |> 
    # replace labels where we're not confident or there are multiplets
    mutate(our_annotation = case_when(
        seurat_clusters %in% c(7, 13) ~ "Multiplets",
        percent < 60 ~ "Not confident", 
        TRUE ~ final_level_labels)) |> 
    rename(cluster_full = seurat_clusters) |> 
    select(cluster_full, our_annotation)

# create a metadata data frame with our new cluster annotations
# (merge with current metadata to get cell identifiers)
new_annotation <- projected@meta.data |> 
    rownames_to_column() |> 
    inner_join(cell_annotations, by = "cluster_full")
rownames(new_annotation) <- new_annotation$rowname    
new_annotation <- subset(new_annotation, select = -rowname)

# add our_annotations into the metadata for projected 
projected <- AddMetaData(projected, new_annotation)

# make the default Ident be "our_annotation"
# otherwise it gives a bunch of problems
Idents(projected) <- "our_annotation"

# save projected data
write_rds(projected, "data/rds_files/05-projected.rds")

# JoinLayers and save as a separate object 
write_rds(JoinLayers(projected), "data/rds_files/06-projected_joined.rds")
```




```{r}


# THIS CHUNK MAPS EACH CELL CLUSTER TO AN ANNOTATION BASED ON AZIMUTH AND OUR
# OBSERVATIONS

# this time we are annotating based on the azimuth_medium category to start, rather than another one that was more detailed. 

# describe each cluster with the most likely azimuth name
cell_annotations <- for_azimuth@meta.data |> 
   
    # count number of cells with each label in each cluster
    group_by(seurat_clusters, azimuth_medium) |> 
    count() |> 
    # for each cluster, use the most common label and count 
    # the percent of cells with that label
    group_by(seurat_clusters) |> 
    arrange(desc(n)) |>
    mutate(percent = n/sum(n)*100) |> 
    slice(1) |> 
    arrange(percent) |> 
    # replace labels where we're not confident or there are multiplets
    mutate(our_annotation = case_when(
        seurat_clusters %in% c(7, 13) ~ "Multiplets",
        percent < 60 ~ "Not confident", 
        TRUE ~ azimuth_medium)) |> 
    rename(cluster_full = seurat_clusters) |> 
    select(cluster_full, our_annotation)







# create a metadata data frame with our new cluster annotations
# (merge with current metadata to get cell identifiers)
new_annotation <- projected_joined@meta.data |> 
    rownames_to_column() |> 
    inner_join(cell_annotations, by = "cluster_full")
rownames(new_annotation) <- new_annotation$rowname    
new_annotation <- subset(new_annotation, select = -rowname)




projected_joined <- read_rds("data/rds_files/06-projected_joined.rds")


# add our_annotations into the metadata for projected 
projected_joined <- AddMetaData(projected_joined, new_annotation)

# make the default Ident be "our_annotation"
# otherwise it gives a bunch of problems
Idents(projected_joined) <- "our_annotation.y"

#tell seurat that when we join layers we want it to be done on the sketch, not the full dataset (which is default)
DefaultAssay(projected_joined)<-"sketch"


#join the layers within the sketch
projected_joined <-JoinLayers(projected_joined)


#find the top markers characterizing the azimuth medium annotations, in the sketch data
projected_joined_markers <-FindAllMarkers(projected_joined, assay = "sketch")

write_csv(projected_joined_markers, "results/all_markers_azimuth_medium_annotations.csv")
# use the markers df to select the top10 genes by lofFC for each cluster

top10<-projected_joined_markers %>%
    group_by(cluster) %>%
    filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup()


#now make a heat map. Go back to our big seurat object which contains our expression levels and plot those top10 genes for each cluster that we discovered above. Expect 10 genes with higher expresion than others for each cluster. 
DoHeatmap(projected_joined, features = top10$gene) +NoLegend()

#what genes were make up the "T cell" group? We see that there are markers for different types of T cells within this group (expected), like IL17, CTLA4, ICOS, various levels of the different CD3's....
projected_joined_markers %>%
    group_by(cluster) %>%
    filter(avg_log2FC > 1,
        cluster == "T cell") %>%
    arrange(p_val_adj) %>% View()
    ungroup()





```

## Plot projected data vs sketch and new annotations

```{r}
# THIS CHUNK CREATES PLOTS COMPARING THE SKETCH WITH THE FULL DATA AND THE 
# NEW ANNOTATIONS

p1 <- DimPlot(projected, label = T, label.size = 3, reduction = 'umap', group.by = 'seurat_clusters') + NoLegend()
p2 <- DimPlot(projected, label = T, label.size = 3, reduction = 'full.umap', group.by = 'cluster_full', alpha = 0.1) + NoLegend()

p1 + p2

ggsave("images/umap_sketch_vs_all_cells.png", width = 10, height = 6, dpi = 600)

# color umap with new annotations
DimPlot(projected, label = T, label.size = 3, reduction = 'full.umap', 
    group.by = 'our_annotation', alpha = 0.1) + 
    NoLegend()

VlnPlot(projected, features = c("CD3D"), group.by = "our_annotation", pt.size = 0)
DotPlot(projected, features = c("CD3D", "CD4", "CD8A", "NKG7", "CD14", "CD19"))
```

```{r read in processed sketch}
projected_joined <- read_rds("data/rds_files/06-projected_joined.rds")
```

## Top markers for each population

```{r}
# THIS CHUNK CALCULATES THE TOP MARKERS FOR EACH CELL POPULATION

# FindAllMarkers takes a long time to run on the full data set in the 
#       "projected_joined" Seurat object
# To speed up, run it only on the sketch cells (contained within 
#       "projected_joined"). 

# Doing so requires the following steps
# 1. Change the default assay from "RNA" to "sketch" to tell Seurat to use the 
#       data stored for the sketch cells
# 2. Subset the Seurat object to only include the sketch cells (I'm not sure 
#       this is actually necessary)
# 3. Redo JoinLayers (we already ran this command on the full projected dataset 
#       but for some reason you have to run it again with the DefaultAssay set
#       to "sketch")

# Note: we have to start with the Seurat object "projected_joined" because that's
#       where we added the cell annotations. The Seurat object "sketch" doesn't
#       contain the cell annotations, so we can't use it

# change default assay to sketch (step 1 above)
DefaultAssay(projected_joined) <- "sketch"

all_markers <- projected_joined |> 
    # get just the cells in the sketch to speed up the code 
    # these cells are identified by having a value in the "seurat_clusters"
    # metadata column, which was only calculated for the sketch cells
    # (step 2 above)
    subset(subset = !is.na(seurat_clusters)) |> 
    # redo JoinLayers on the sketch assay (step 3 above)
    JoinLayers() |> 
    FindAllMarkers(only.pos = TRUE)

write_csv(all_markers, "results/all_markers.csv")

# change default assay back to RNA (all cells instead of just sketch cells)
# for future use
DefaultAssay(projected_joined) <- "RNA"

# Next step: use top markers to populate DotPlot
```

## Differential expression

```{r}
# THIS CHUNK CALCULATES DE GENES BY CYCLE PHASE

# function to test for differentially expressed genes by phase within specified
# cell population, adjusting for ptid 
do_de_test <- function(data, annotation) {
    print(paste("running", annotation))
    # subset the seurat object to cells of interest
    subsetted <- subset(data, subset = our_annotation.y == annotation) 
    
    # incantation that stops using BPCells and puts data into memory
    # necessary for running any FindMarkers test other than wilcox
    mat <- LayerData(subsetted, assay = "RNA", layer = "counts")
    mat <- as(mat, "dgCMatrix")
    subsetted[["RNA"]] <- CreateAssayObject(counts = mat)
    
    # delete the matrix object to save memory 
    # (it has already been saved inside the Seurat object so is okay to delete)
    rm(mat)
    
    # specify comparison between phases
    Idents(subsetted) <- "Phase"
    
    de_results <- FindMarkers(subsetted, 
        ident.1 = "F",
        ident.2 = "L",
        test.use = "MAST", 
        # accounts for pairing between samples
        latent.var = "Ptid") |> 
        # store gene names in a column
        rownames_to_column("gene") |> 
        # record cell population
        mutate(our_annotation = annotation) 
    
    # save a file with the results 
    write_csv(de_results, paste0("results/", annotation, "_de_genes.csv"))
    
    # return the results
    de_results
}

# testing
# x <- subset(projected, downsample = 100)
# x_joined <- JoinLayers(x)
# test <- do_de_test(x_joined, populations[8])

projected_joined <- read_rds("data/rds_files/06-projected_joined.rds")

# get names of each population in reverse order of population size
# so it runs the smallest populations first
populations <- projected_joined@meta.data |> group_by(our_annotation) |> count() |> arrange(n) |> pull(our_annotation)

# run DE tests for every cell population
de_genes_by_phase <- map(populations, function(popn) do_de_test(projected_joined, popn)) |> 
    bind_rows()

# plots to gut check the de_genes results
# VlnPlot(x_joined |> subset(our_annotation == "CD14 monocyte"), "CALR", group.by = "Phase")
# VlnPlot(x_joined |> subset(our_annotation == "CD14 monocyte"), "CALR", group.by = "Ptid", 
#     split.by = "Phase")
# VlnPlot(x_joined |> subset(our_annotation == "CD16 NK cell"), "SPPL2B", group.by = "Ptid", 
#     split.by = "Phase")

Idents(projected_joined) <- "Sample"
to_plot <- projected_joined |> subset(downsample = 1000)

de_genes <- list.files("results", full.names = TRUE, pattern = "de_genes") |> 
    read_csv() |> 
    bind_rows()



VlnPlot(projected_joined |> subset(our_annotation == "CD8 T cell"), "HBB", 
    group.by = "Ptid", split.by = "Phase")






# function to test for differentially expressed genes by phase within specified
# cell population, adjusting for ptid 
do_de_test_pseudobulked <- function(data, annotation) {
    print(paste("running", annotation))
    # subset the seurat object to cells of interest
    subsetted <- subset(data, subset = our_annotation.y == annotation) |> 
        AggregateExpression(return.seurat = TRUE, group.by = c("Ptid", "Phase"))
    
    # specify comparison between phases
    Idents(subsetted) <- "Phase"
    
    de_results <- FindMarkers(subsetted, 
        ident.1 = "F",
        ident.2 = "L",
        test.use = "DESeq2") |> 
        # store gene names in a column
        rownames_to_column("gene") |> 
        # record cell population
        mutate(our_annotation = annotation) 
    
    # save a file with the results 
    write_csv(de_results, paste0("results/", annotation, "_pseudobulked_de_genes.csv"))
    
    # return the results
    de_results
}



# get names of each population in reverse order of population size
# so it runs the smallest populations first
populations <- projected_joined@meta.data |> group_by(our_annotation.y) |> count() |> arrange(n) |> pull(our_annotation.y)


DefaultAssay(projected_joined) <- "RNA"
# run DE tests for every cell population PSEUDOBULKED
de_genes_by_phase_pseudobulked <- map(populations, function(popn) do_de_test_pseudobulked(projected_joined, popn)) |> 
    bind_rows()

# run DE tests for every cell population MAST LATENT VAR
de_genes_by_phase <- map(populations, function(popn) do_de_test(projected_joined, popn)) |> 
    bind_rows()


do_de_test_pseudobulked(projected_joined, "CD8 T cell")


cd8_DE <-read_csv("results/CD8 T cell_pseudobulked_de_genes.csv")

projected_joined@meta.data$our_annotation %>% unique()


ggplot(cd8_DE, aes(x = avg_log2FC, y = -log10(p_val)))+
    geom_point()+
    geom_hline(yintercept = -log10(0.05))




cd8_DE %>%
    filter(abs(avg_log2FC) >= 0.2,
        p_val <= 0.05) %>% View()




VlnPlot(projected_joined |> subset(our_annotation == "HSC"), "GSPT1", 
    group.by = "Ptid", split.by = "Phase")





```



TODO: 

* Run FindAllMarkers on full dataset to identify top genes associated with each annotation
    * Make nice dot plot justifying annotations based on top markers
* Run DE on full projected data set
* Plot number of cells in each cell type vs phase (i.e. are there more CD4 T cells during follicular than luteal?)
* Switch to Hyak

```{r}
sessionInfo()
```